---
title: "Homework2"
author: "Ethan Ashby"
date: "2/10/2020"
output: pdf_document
---

```{r Packages, message=FALSE, echo=FALSE}
library(mosaic)
library(tidyverse)
```
1. Explore lab 1 data

(a) Write down model that you think works.
```{r}
labdata<-read.csv("Lab1data.txt")
summary(labdata)

labdata$sj <- with(labdata, sqrt(Jobs))
xyplot(Time~sj, data=labdata, type=c('r', 'p'))
#the sqrt transformation on 'Jobs' improves the linear fit! I feel comfortable applying the model here
```

```{r}
#sqrt transform jobs data
labdata$sj <- with(labdata, sqrt(Jobs))

#fit lm to this data, with y intercept set to 0, since 0 jobs should take 0 minutes
fit<-lm(Time~0+sj, labdata)

summary(fit)
```

\begin{equation}
time=3.979*\sqrt{jobs}
\end{equation}

```{r}
#Time required to do 10 jobs given our model
predict(fit, data.frame(sj=sqrt(10)))
```

(c) In the context of this data, make sense of the form the regression relationship might take. In particular, discuss why we might see a linear relationship, a convex relationship, or the concave relationship seen in the actual data. Keep this specific to the variables we are studying.

```{r}
#relationship in the actual data
xyplot(Time~Jobs, labdata)
```

We might anticipate seeing a concave relationship (which we do see in the data) because a small number of jobs requires a mininum amount of time to complete. However, when posed with a large number of jobs, the process becomes more efficient and it takes less time to complete each job, resulting in a tapering off of the data at higher number of jobs. This accounts for the concavity of the data.

If the relationship is convex, then as the number of jobs increases, the process becomes less efficient and the time/job increases. So at a high number of jobs, the time required to complete explodes.

If the relationship is linear, then as the time/job is constant regardless of how many jobs we have.

2. Based on the work we did understanding the forms of the variances of the of the regression coefficients, discuss (confidence) interval estimation for a quantity like the one in 1(b). Should the size of the interval be a function of $x^*$? If so, how and why? Where do you think the smallest interval would be for a fixed confidence level?

```{r}
#use predict.lm with argument 'interval="confidence"' to generate confidence interval for mean time at 10 jobs
summary(fit)

#Time to complete 10 jobs
predict.lm(fit, data.frame(sj=sqrt(10)), interval="confidence") 

#Time to complete 100000 jobs
predict.lm(fit, data.frame(sj=sqrt(100000)), interval="confidence") 
```

As evidenced by comparing confidence interval predictions for 10 and 100,000 jobs, the larger number of jobs results in a wider interval. So the size of the interval increases as the number of jobs increases (interval width is a function of x*).
I would anticipate the smallest interval for a fixed confidence level would be the origin.

3. Show that the regression line necessarily passes through the point $(\bar{x},\bar{y})$.

We know that this is true because least squares regression defines $$\hat{\beta_0}=\bar{y}-\hat{\beta_1}\bar{x}$$. A rearrangement of this expression shows that $$\hat{\beta_0}+\hat{\beta_1}\bar{x}=\bar{y}$$, which demonstrates that $(\bar{x},\bar{y})$ satisfy the equation of the line of best fit. Thus the line must pass through this point.

4. Consider a situation where the x variables are measured with error, so that rather than observing (xi, yi) we observe $(x_i + \eta_i, y_i)$ where the $y_i$ are the responses at the actual $x_i$ that we don’t get to see and $\eta_i$ is some mean 0 random variable (similar to $\epsilon_i$). The consequence of this is that the least squares estimate of the slope will be biased. Using intuition, argue which direction the bias will be (negative bias: $\hat{\beta}_1$ will tend to be too small, vs positive bias).

Recall that $$\hat{\beta}_1=\frac{\sum{(x_i-\bar{x})(y_i)}}{\sum{(x_i-\bar{x})^2}}$$

Jittering the x values slightly in your measurement will not change the value of $\bar{x}$, since the $\eta_i$ is normally distributed with mean=0. So some $x_i$ will be positively perturbed, and some will be negatively perturbed, so the $\bar{x}$ will remain the same. Jittering the x values also won't change the value of $\sum{(x_i-\bar{x})(y_i)}$, since $\sum{((x_i+\eta_i)-\bar{x})(y_i)}=\sum{(x_i-\bar{x})(y_i)}+\sum{(\eta_i)(y_i)}$ and $\sum{(\eta_i)(y_i)}=0$ because the jitter is independent of the value of the response variable. $\sum{((x_i+\eta_i)-\bar{x})(y_i)}=\sum{(x_i-\bar{x})(y_i)}$, so the jitter has no effect on the numerator of the $\hat{\beta}_1$ expression.

However, the denominator of the expression for $\hat{\beta}_1$ will increase.
Compare the unperturbed sum ($\sum(x_i-\bar{x})^2$) and the perturbed sum ($\sum((x_i+\eta_i)-\bar{x})^2$).
$\sum(x_i-\bar{x})^2=\sum({x_i}^2-2\bar{x}x_i+\bar{x}^2)$ and $\sum((x_i+\eta_i)-\bar{x})^2=\sum {x_i}^2-2x_i\bar{x}+2x_in_i+\bar{x}^2-2\bar{x}n_i+{n_i}^2$.
These expressions have a lot of shared terms. To determine which is larger, we can cancel many of these shared terms (${x_i}^2$, $2x_i\bar{x}$, and $\bar{x}^2$). Then the perturbed equation remains as $\sum 2x_i\eta_i-2\bar{x}\eta_i+\eta_i^2$.
Note that $\sum 2x_i\eta_i-2\bar{x}\eta_i=0$ because $\sum x_i-\bar{x}=0$. So this will simplify to $\sum \eta_i^2$ which is necessarily greater than 0.
So $\sum(x_i-\bar{x})^2 < \sum((x_i+\eta_i)-\bar{x})^2$.
So the numerator is unchanged by the jittering, but the denominator is larger. So the expression for $\hat{\beta}_1$ will have an unchanged numerator and a larger denominator, resulting in an underestimation of $\hat{\beta}_1$, or a negative bias.









