---
title: "Homework 1"
author: "Ethan Ashby"
date: "1/29/2020"
output: pdf_document
---

```{r}
library(tidyverse)
```


1. Consider observing n = 10 pairs of points where no clear predictor/response
relationship is obvious. Supposing that the relationship between the two variables appears to be linear, find the range of r (correlation coefficient) for which
the null hypothesis H0 : ρ = 0 will not be rejected at α = .05. (Refer to the
t-test given at the end of lecture notes 2). (If you want to get quantiles from
a t-distribution in R, you can use qt(quantile,df ) where quantile is a number
between 0 and 1 and df is the degrees of freedom.)

6.  Using the data set mtcars (data(mtcars)) in R, find a transformation of the
variables mpg (fuel efficiency) and disp (engine size) that allows for reasonable
application of our linear model.

```{r}
data(mtcars)

#diagnostic plot
mtcars %>% select(mpg, disp) %>% ggplot()+geom_point(aes(x=mpg, y=disp))+ggtitle("Untransformed mpg vs disp from mtcars")

#try a log transform on disp variable
mtcars %>% select(mpg, disp) %>% ggplot()+geom_point(aes(x=mpg, y=log(disp)))+ggtitle("mpg vs Log transformed disp from mtcars")

#The data looks quite a bit more linear now... this is more reasonable for applying our linear model
```

